### 2024-03-12 manuscript.tex
Italicized gene name
Added middle initials to authors
Removed conflicts of interest and funding sections because entered elsewhere in submission
Added author summary

Marc Singleton\textsuperscript{1,*}, Michael Eisen\textsuperscript{1,2}

@@

@@

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig~\ref{fig:alignment}D), whereas the second does not (Fig~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene {ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains~\cite{Mitchell1989, Gerber1994}, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features. In contrast, the low log likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant log likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.}, columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining this information with a protein's biological context to jointly suggest testable hypotheses of function.

@@

\section{Funding}
This work was supported by the Howard Hughes Medical Institute.

\section{Conflicts of interest}
The authors declare no conflicts of interest.

>>>>

Marc D. Singleton\textsuperscript{1,*}, Michael B. Eisen\textsuperscript{1,2}

@@

\section{Author summary}
Proteins are the molecular machines that carry out many processes required for life at an atomic level. Though many proteins use fixed structures to perform their functions, proteins with flexible segments are widespread, especially in multicellular organisms. Furthermore, these intrinsically disordered regions (IDRs) are often involved in essential cellular functions. However, the sequences of IDRs evolve quickly, which challenges traditional bioinformatics methods that depend on sequence conservation to predict function. Several studies have demonstrated that distributed biophysical features of IDRs are constrained rather than their exact sequences, and a recent study in yeast found that IDRs with common patterns of conserved features were associated with specific functions. Therefore, in this work we ask if IDRs in fruit flies, another common laboratory organism, also have patterns of conservation with associated functions. We build on the previous study by integrating their approach into a fully statistical framework based on mathematical models of trait evolution. Though we identify widespread signals of conservation in the IDRs of fruit flies, we find less evidence of a simple relationship between features and function. These methods and results will provide a valuable resource that can guide future experimental analyses of IDRs in fruit flies and other organisms.

@@

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig~\ref{fig:alignment}D), whereas the second does not (Fig~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene \textit{ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains~\cite{Mitchell1989, Gerber1994}, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features. In contrast, the low log likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant log likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.}, columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining this information with a protein's biological context to jointly suggest testable hypotheses of function.

@@
